---
title: "Quick tour through Markdown"
author: "Anaga and Aaron"
date: "2025-04-15"
output:
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

-   Already installed in all R Studio (File \> New File \> R Markdown)
-   Markdown makes it easier for users to write code and free-text at the same time
-   Code is organized into "chunks" that can be run individually
-   Markdown files can be used to create (called **"knit"**) files
in HTML, PDF, Word, Powerpoint, and more
-   If your end product is a report of statistics, Markdown can be especially useful to automate making a report
-   Markdown is compatible with Github and traditional R language as well as Python, Bash, SQL, and more

## Let's download knitr

```{r, echo=T, warning=F}
#install.packages("knitr")
library(knitr)
```

## In-line expressions

With Markdown you can write sentences that reference elements of your data analysis.

For example, let's say we wanted to describe the chicken weights dataset. How many chickens are there?

```{r cars, echo = TRUE}
data(chickwts)
head(chickwts, n=10)
nrow(chickwts)
```

In-line expressions are written as backtick (`) r, then some code, and then another backtick. (Demonstrate)

These are the same: 

-   "There were 71 chickens" (the explicit approach)
-   "There were `r nrow(chickwts)` chickens" (the in-line approach) 

This can also be helpful: 

- The mean weight of the chickens was `r round(mean(chickwts$weight),2)` units. 

```{r, echo=T}
mod <- summary(lm(weight ~ feed, data=chickwts))
```

- Compared to casein feed, horsebean feed was associated with `r mod$coefficients[2]` units difference weight.

## Contingency tables

If you're creating many, many two-way tables and want an efficient way of creating a report that contains your tables, Markdown is very helpful.

```{r, echo=T}
chickwts$dead <- sample(c(0,1), size=nrow(chickwts), replace = T)
chickwts$color <- sample(c("blue", "green", "red"), size=nrow(chickwts), replace=T)
chickwts$eyecolor <- sample(c("yellow", "orange", "purple"), size=nrow(chickwts), replace=T)

table(chickwts$dead, chickwts$feed, chickwts$color)
table(chickwts$dead, chickwts$feed, chickwts$eyecolor)

```

## Tips and tricks to get started with Markdown

-   You can toggle between "Source" and "Visual" to preview your script as an HTML product (can help review your own work). The source-visual toggle is located below the save button.
-   You can use hashtags to create headers and create sections of your report. Open Outline (on the upper right hand side) and navigate your own document by section.
- Each chunk comes with a play button (to run the chunk), play all prior chunks, and a settings button. The settings allow you to control how the chunk will appear and work in the final document. For example, you can control when to show/hide outputs (**"echo"**), show/hide warnings (**"warning"**), show/hide messages (**"message"**).
- As long as you have datasets in your working environment (right), you can use Markdown alongside traditional R scripts. Try using **source()** at the top of your Markdown file, and writing comments from your analysis in Markdown.

## Plots

Yes, you can also include plots in Markdown. Note: You may still need to adjust the size of your plots according to your output file.

```{r plot, echo=T, warning=F, message=F}
# note- this chunk has warning=F and message=F to hide the messages from uploading these packages!
library(ggplot2)
library(tidyverse)
chickwts %>%
  ggplot(aes(x = feed, y = weight)) + 
  geom_boxplot() + 
  geom_point()

chickwts %>%
  ggplot(aes(x = feed, y = weight)) + 
  geom_boxplot() + 
  geom_point() +
  facet_wrap(~dead)
```


## YouTube videos and other helpful resources

-   <https://www.youtube.com/watch?v=DNS7i2m4sB0>
-   <https://www.youtube.com/watch?v=asHhuHRxhvo>
- <https://rmarkdown.rstudio.com/index.html> 
