---
title: "Quick tour through Markdown"
author: "Anaga and Aaron"
date: "2025-04-15"
output:
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Benefits of R Markdown

-   You can code and include free text.
-   You can create reports in HTML, PDF, Word, Powerpoint, and more!
-   You can make updates to reports more efficiently.
-   You can use Markdown along with R scripts.

## Let's download knitr

```{r, echo=F, warning=F}
#install.packages("knitr")
library(knitr)
```

## You can use Markdown to reference your code in the text

Can you reference chunks in text? I had excluded people from the dataset (see chunk cars).

```{r cars, echo = TRUE}
data(chickwts)
head(chickwts, n=10)
nrow(chickwts)
```

## Slide with Plot

When you're working in markdown it can be hard to keep track of your writing especially when you're writing long notes like this. i am aaron. i talk a lot. no i don't. this is my inner dialogue. banana.

```{r plot, message=F}
library(ggplot2)
library(tidyverse)
chickwts %>%
  ggplot(aes(x = feed, y = weight)) + 
  geom_boxplot() + 
  geom_point()
```

## YouTube videos

-   <https://www.youtube.com/watch?v=DNS7i2m4sB0>
-   <https://www.youtube.com/watch?v=asHhuHRxhvo>
